---
title: ""
runtime: 
toc: false
---

```{r}
#| include: false
knitr::opts_chunk$set(
  echo = FALSE, #hide code
  message = FALSE, #hide package messages
  warning = FALSE #hide warnings
)

```

::: country-hero-text
<h1 class="country-title">

</h1>

<h2>Benin Health Facilities</h2>

<p>Standardized, geolocated facility information to strengthen modelling, planning, and decision-making</p>
:::

<!-- ===================== -->

<!-- Map + Section Navigation -->

<!-- ===================== -->

::::: {.columns .country-main}
::: {.column width="70%"}
## Health Facilities Map

```{r}
library(sf) 
library(leaflet)
library(shiny)
library(dplyr)

# Example: load pre processed sf object

hf_sf <- readRDS("../data/Benin/benin_adm2_monthly_allages_with_rainfall.rds")

leaflet(hf_sf |> sf::st_centroid()) |>
  addTiles() |> 
  addCircleMarkers(radius = 3, stroke = FALSE, fillOpacity = 0.7)

```
:::

::: {.column .toc-panel width="30%"}
### On this page

-   [Overview](#overview)
-   [Demographics](#demographics)
-   [Services](#services)
-   [Data Download](#data-download)
:::
:::::

## Overview {#overview}

Benin's health system is structured into three levels: peripheral (health zones), intermediate (departmental), and central (national). The country has around 900 public health facilities complemented by a growing private sector. Health zone offices coordinate service delivery across Benin's 12 departments. Despite improvements, access to care remains unequal, with the northern regions facing greater barriers due to distance and infrastructure.

## Demographics {#demographics}

Benin has a population of approximately 13 million, with high fertility and a young population structure. Most public facilities are community health centres (CSC/CCS) serving rural communities. The country has a significant shortage of skilled health personnel, particularly in rural zones, with a physician density of approximately 0.7 per 10,000 population. Community health agents (relais communautaires) extend the reach of facility-based care.

## Services {#services}

Health facilities in Benin provide outpatient consultations, antenatal care, malaria prevention (IPTp, ITN distribution) and treatment, HIV testing and ART, immunization under the Expanded Programme on Immunization (EPI), nutrition services, and family planning. Malaria is the leading cause of morbidity and mortality in Benin, particularly in children under five, with high transmission in the south.

## Data Download {#data-download}

Download the facilities dataset here:

-   [CSV](../data/Benin/hf_demographics.csv)\
<!--   [JSON](../data/Benin/hf_demographics.json)\ -->
-   [RDS](../data/Benin/benin_adm2_monthly_allages_with_rainfall.rds)
