---
title: ""
runtime:
toc: false
---

```{r}
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

::: country-hero-text
<h1 class="country-title">

</h1>

<h2>Ethiopia Health Facilities</h2>

<p>Standardized, geolocated facility information to strengthen modelling, planning, and decision-making</p>
:::

::::: {.columns .country-main}
::: {.column width="70%"}
## Health Facilities Map

```{r}
library(sf)
library(leaflet)
library(dplyr)

hf_sf <- read.csv("../etl/data/processed/country_standardized/ethiopia_standardized.csv") |>
  dplyr::select(facility_name, facility_type, latitude, longitude, admin1) |>
  dplyr::mutate(
    latitude  = as.numeric(latitude),
    longitude = as.numeric(longitude)
  ) |>
  dplyr::filter(
    !is.na(latitude) & !is.na(longitude),
    latitude  >= 3   & latitude  <= 15,
    longitude >= 33  & longitude <= 48
  ) |>
  sf::st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

leaflet(hf_sf) |>
  addTiles() |>
  addCircleMarkers(radius = 2, stroke = FALSE, fillOpacity = 0.6)
```
:::

::: {.column .toc-panel width="30%"}
### On this page

-   [Overview](#overview)
-   [Demographics](#demographics)
-   [Services](#services)
-   [Data Download](#data-download)
:::
:::::

## Overview {#overview}

Ethiopia operates one of the largest health systems in sub-Saharan Africa, organised in a three-tier structure: primary health care (health posts, health centres and primary hospitals), general hospitals, and specialised / tertiary hospitals. The Federal Ministry of Health maintains a centralised e-HFR (electronic Health Facility Registry), with this dataset reflecting a recent extract spanning all 11 regional administrations and city administrations.

## Demographics {#demographics}

With a population of approximately 125 million — the second-largest in Africa — Ethiopia faces enormous service-delivery demand. The Health Extension Programme has placed two trained extension workers at each rural kebele-level health post, making the country a continental reference for community-based primary care. Doctor density remains low (~1.1 per 10,000), with nurses, midwives and Health Extension Workers carrying most of the front-line workload.

## Services {#services}

The e-HFR records four service domains per facility — clinical, diagnostic, support, and 24-hour services — which the HealthScope ETL pipeline unnests and maps to the standard 3-level hierarchy used for every other country. Coordinates were joined from the [Humanitarian Data Exchange Ethiopia Health Facilities dataset](https://data.humdata.org/dataset/ethiopian-health-facilities) (~63% direct match by Facility ID); facilities without an HDX coordinate (mostly drug stores and small private clinics) are excluded from the map.

```{r}
library(dplyr)
library(ggplot2)
library(scales)

svc <- read.csv("../etl/data/processed/country_standardized/ethiopia_services_standardized.csv",
                stringsAsFactors = FALSE) |>
  dplyr::filter(include_in_analysis == "TRUE" | include_in_analysis == TRUE)

n_facilities <- length(unique(svc$facility_code))

top_services <- svc |>
  dplyr::distinct(facility_code, service_name) |>
  dplyr::count(service_name, name = "facilities") |>
  dplyr::mutate(pct = facilities / n_facilities) |>
  dplyr::arrange(dplyr::desc(facilities)) |>
  utils::head(15)

ggplot2::ggplot(top_services,
                ggplot2::aes(x = facilities,
                             y = stats::reorder(service_name, facilities))) +
  ggplot2::geom_col(fill = "#c9a227") +
  ggplot2::geom_text(ggplot2::aes(label = scales::percent(pct, accuracy = 1)),
                     hjust = -0.15, size = 3.2, colour = "#3d3d3d") +
  ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0, 0.18))) +
  ggplot2::labs(
    title = "Top 15 services across Ethiopian facilities",
    subtitle = paste0("Share of ", scales::comma(n_facilities), " facilities offering each service"),
    x = "Facilities offering service",
    y = NULL
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_blank(),
    panel.grid.minor   = ggplot2::element_blank(),
    plot.title    = ggplot2::element_text(face = "bold", colour = "#3d3d3d"),
    plot.subtitle = ggplot2::element_text(colour = "#6b6b6b")
  )
```

```{r}
domain_coverage <- svc |>
  dplyr::distinct(facility_code, service_domain) |>
  dplyr::count(service_domain, name = "facilities") |>
  dplyr::mutate(pct = facilities / n_facilities) |>
  dplyr::arrange(dplyr::desc(facilities))

ggplot2::ggplot(domain_coverage,
                ggplot2::aes(x = facilities,
                             y = stats::reorder(service_domain, facilities))) +
  ggplot2::geom_col(fill = "#5a7a9b") +
  ggplot2::geom_text(ggplot2::aes(label = scales::percent(pct, accuracy = 1)),
                     hjust = -0.15, size = 3.2, colour = "#3d3d3d") +
  ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0, 0.18))) +
  ggplot2::labs(
    title = "Facility coverage by service domain",
    subtitle = "Share of facilities offering at least one service in each domain",
    x = "Facilities",
    y = NULL
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_blank(),
    panel.grid.minor   = ggplot2::element_blank(),
    plot.title    = ggplot2::element_text(face = "bold", colour = "#3d3d3d"),
    plot.subtitle = ggplot2::element_text(colour = "#6b6b6b")
  )
```

## Data Download {#data-download}

Download Ethiopia datasets here:

-   Facilities &mdash; [CSV](../etl/data/processed/country_standardized/ethiopia_standardized.csv)
-   Services &mdash; [CSV](../etl/data/processed/country_standardized/ethiopia_services_standardized.csv)

Produced by the [HealthScope ETL pipeline](https://github.com/map-data-engineering/health_facility_etl) from the Ethiopia MoH e-HFR, with coordinates joined from HDX.
