---
title: ""
runtime:
toc: false
---

```{r}
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

::: country-hero-text
<h1 class="country-title">

</h1>

<h2>Ethiopia Health Facilities</h2>

<p>Standardized, geolocated facility information to strengthen modelling, planning, and decision-making</p>
:::

::::: {.columns .country-main}
::: {.column width="70%"}
## Health Facilities Map

```{r}
library(leaflet)
library(data.table)

# data.table::fread + select keeps memory low on the 40K-row
# Ethiopia file; leaflet accepts plain data frames with lng/lat
# columns, no sf conversion needed.
hf <- data.table::fread(
  "../etl/data/processed/country_standardized/ethiopia_standardized.csv",
  select = c("facility_name", "facility_type", "latitude",
             "longitude", "admin1")
)
hf <- hf[!is.na(latitude) & !is.na(longitude) &
         latitude  >= 3  & latitude  <= 15 &
         longitude >= 33 & longitude <= 48]

leaflet(hf) |>
  addTiles() |>
  addCircleMarkers(lng = ~longitude, lat = ~latitude,
                   radius = 2, stroke = FALSE, fillOpacity = 0.6)
```
:::

::: {.column .toc-panel width="30%"}
### On this page

-   [Overview](#overview)
-   [Demographics](#demographics)
-   [Services](#services)
-   [Data Download](#data-download)
:::
:::::

## Overview {#overview}

Ethiopia operates one of the largest health systems in sub-Saharan Africa, organised in a three-tier structure: primary health care (health posts, health centres and primary hospitals), general hospitals, and specialised / tertiary hospitals. The Federal Ministry of Health maintains the centralised [Master Facility Registry v2](https://mfrv2.moh.gov.et/#/facility/all), and this dataset reflects a recent CSV export from that registry covering all 11 regional administrations and the two chartered city administrations.

## Demographics {#demographics}

With a population of approximately 125 million — the second-largest in Africa — Ethiopia faces enormous service-delivery demand. The Health Extension Programme has placed two trained extension workers at each rural kebele-level health post, making the country a continental reference for community-based primary care. Doctor density remains low (~1.1 per 10,000), with nurses, midwives and Health Extension Workers carrying most of the front-line workload.

## Services {#services}

MFR v2 records four service domains per facility — clinical, diagnostic, support, and 24-hour services — which the HealthScope ETL pipeline unnests and maps to the standard 3-level hierarchy used for every other country. The public MFR v2 CSV export does not include GPS, so coordinates are joined from the openly-licensed WHO AFRO compilation on the [Humanitarian Data Exchange](https://data.humdata.org/dataset/ethiopian-health-facilities) (~63% direct match by Facility ID); facilities without an HDX coordinate (mostly drug stores and small private clinics outside WHO AFRO's traditional coverage scope) are excluded from the map.

```{r}
library(dplyr)
library(ggplot2)
library(scales)

# Pre-aggregated summary — the full Ethiopia services CSV (84 MB)
# is too large to load in this render. Regenerate with
# scripts/aggregate_ethiopia_stats.R after a fresh ETL refresh.
.eth_summary <- readRDS("../data/Ethiopia/ethiopia_services_summary.rds")
n_facilities    <- .eth_summary$n_facilities
top_services    <- .eth_summary$top_services
domain_coverage <- .eth_summary$domain_coverage

ggplot2::ggplot(top_services,
                ggplot2::aes(x = facilities,
                             y = stats::reorder(service_name, facilities))) +
  ggplot2::geom_col(fill = "#c9a227") +
  ggplot2::geom_text(ggplot2::aes(label = scales::percent(pct, accuracy = 1)),
                     hjust = -0.15, size = 3.2, colour = "#3d3d3d") +
  ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0, 0.18))) +
  ggplot2::labs(
    title = "Top 15 services across Ethiopian facilities",
    subtitle = paste0("Share of ", scales::comma(n_facilities), " facilities offering each service"),
    x = "Facilities offering service",
    y = NULL
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_blank(),
    panel.grid.minor   = ggplot2::element_blank(),
    plot.title    = ggplot2::element_text(face = "bold", colour = "#3d3d3d"),
    plot.subtitle = ggplot2::element_text(colour = "#6b6b6b")
  )
```

```{r}
ggplot2::ggplot(domain_coverage,
                ggplot2::aes(x = facilities,
                             y = stats::reorder(service_domain, facilities))) +
  ggplot2::geom_col(fill = "#5a7a9b") +
  ggplot2::geom_text(ggplot2::aes(label = scales::percent(pct, accuracy = 1)),
                     hjust = -0.15, size = 3.2, colour = "#3d3d3d") +
  ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0, 0.18))) +
  ggplot2::labs(
    title = "Facility coverage by service domain",
    subtitle = "Share of facilities offering at least one service in each domain",
    x = "Facilities",
    y = NULL
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_blank(),
    panel.grid.minor   = ggplot2::element_blank(),
    plot.title    = ggplot2::element_text(face = "bold", colour = "#3d3d3d"),
    plot.subtitle = ggplot2::element_text(colour = "#6b6b6b")
  )
```

## Data Download {#data-download}

Download Ethiopia datasets here:

-   Facilities &mdash; [CSV](../etl/data/processed/country_standardized/ethiopia_standardized.csv)
-   Services &mdash; [CSV](../etl/data/processed/country_standardized/ethiopia_services_standardized.csv)

Produced by the [HealthScope ETL pipeline](https://github.com/map-data-engineering/health_facility_etl) from the [Ethiopia MoH Master Facility Registry v2](https://mfrv2.moh.gov.et/#/facility/all), with coordinates joined from the WHO AFRO compilation on HDX.
